Brother, Can You Spare a Manufacturing Job? How Voters React to Deindustrialization

### 2021-12-12 Final Draft/Final Draft.Rmd
pushing changes to final draft

This paper analyzes the relationship between deindustrialization, race, and the election of Donald Trump in 2016. Previous analysis has frequently focused on racial animus, trade, or the conjunction of the two (\cite{Autor20}; \cite{Che16}; \cite{BR21}). We extend the analysis of \cite{Baccini21}, one of the first papers to examine the political effects of deindustrialization while considering race and localization. Deindustrialization - the transition of the U.S. economy away from manufacturing - has led to widespread job loss and created adverse knock-on effects for a wide swath of Americans. This incorporates job losses due to trade with China and others (see \cite{Acemoglu16}); it also includes job losses due to automation, demand changes, and other factors. The experience of deindustrialization has not been uniform; rather, it has been concentrated in certain areas of the country that were previously economically reliant on manufacturing, as demonstrated in Figure \ref{ManuMap}. We believe the localized nature of this analysis highlights the uneven and concentrated harms of deindustrialization.

Along with \cite{Baccini21}, we find evidence that the experience of deindustralization was associated with support for Donald Trump, particularly among whites. However, we find that the related effects in the 2016 election are not associated with recent manufacturing job loss; rather, the opposite. Counties with manufacturing job losses from 2012-2015 tended to become more Democratic. We extend the analysis to the 2004-2015 period, to capture the long-term effects of deindustrialization on a region. Doing so, we find that regions experiencing more manufacturing job losses swung towards Donald Trump in 2016, and that this occurred more when those losing jobs were white.

\begin{figure} \label{ManuMap}
\caption{Net Change in Manufacturing Employment from 2004 to 2015}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center"}
# load map
model_input <- readRDS("../model_5_input.rds")

@@

ggplot() +
  geom_sf(data = mapping_county, aes(fill = ntile_mfg),
          color = NA, size = .01) +
  scale_fill_manual(values = colors,
                    labels = labels) +
  # add state lines
  geom_sf(data = state_shape, fill = NA, color = "black", size = .1) +
  coord_sf(datum = NA) +
  theme_minimal() +
  # geom_sf_text_repel(aes(label = state), data = mapping_small,
  #                    force = .00001,  seed = 10,
  #                    size = 2.5) +
  theme(  axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          text = element_text(size = 8)) +
  labs(title = "",
       subtitle = "",
       fill = "") +
  theme(legend.key.size = unit(.4, 'cm'))

@@

@@

\begin{figurenotes}
Gain or loss of national manufacturing jobs, divided by total national workforce.
\end{figurenotes}
\end{figure}

\begin{figure} \label{vShapePlot}
\caption{}

\begin{figurenotes}
Figure notes without optional leadin.
\end{figurenotes}
\end{figure}

@@

\begin{figure} \label{natlPlot}
\caption{Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center"}

@@

>>>>

This paper analyzes the relationship between deindustrialization, race, and the election of Donald Trump in 2016. Previous analysis has  focused on racial animus, trade, or the conjunction of the two on  political polarization and election outcomes (\cite{Autor20}; \cite{Che16}; \cite{BR21}). We extend the analysis of \cite{Baccini21}, one of the first papers to examine the political effects of deindustrialization that while considering race and localization. Deindustrialization - the transition of the U.S. economy away from manufacturing - has led to widespread job loss and created adverse knock-on effects for a wide swath of Americans. Our measure of dindustrialization incorporates job losses due to trade with China and others (see \cite{Acemoglu16}); it also includes job losses due to automation, demand changes, and other factors. The experience of deindustrialization has not been uniform; rather, it has been concentrated in certain areas of the country that were previously economically reliant on manufacturing, as demonstrated in Figure \label{ManuMap}. The localized nature of this analysis highlights the uneven and concentrated harms of deindustrialization.

<!-- FIGURE X: MAP OF DEINDUSTRIALIZATION HERE  -->

\begin{figure} \label{ManuMap}
\caption{Concentration of Deindustrialization (2004-2015) in the eastern Heartland}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center"}
# load map
model_input <- readRDS("../model_5_input.rds")

@@

ggplot() +
  geom_sf(data = mapping_county, aes(fill = ntile_mfg),
          color = NA, size = .01) +
  scale_fill_manual(values = colors,
                    labels = labels) +
  # add state lines
  geom_sf(data = state_shape, fill = NA, color = "black", size = .1) +
  coord_sf(datum = NA) +
  theme_minimal() +
  # geom_sf_text_repel(aes(label = state), data = mapping_small,
  #                    force = .00001,  seed = 10,
  #                    size = 2.5) +
  theme(  axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          text = element_text(size = 8)) +
  labs(title = "Net Change in Manufacturing Employment from 2004 to 2015",
       subtitle = "",
       fill = "") +
  theme(legend.key.size = unit(.4, 'cm'))

@@

Like \cite{Baccini21}, we find evidence that the experience of deindustralization was associated with support for Donald Trump, particularly among whites. \cite{Baccini21}
find that gross manufacturing layoffs between 2012-2015 were associated with a loss in Democratic vote share. However, using a different measure of deindustrialization - *net* change in manufacturing employment - we find the oppposite effect: that counties with manufacturing job *gains* from 2012-2015 tended to swing towards Donald Trump in 2016. We extend the analysis to the 2004-2015 period in order to capture the long-term effects of deindustrialization on a region. Doing so, we find that regions experiencing more manufacturing job losses swung towards Donald Trump in 2016, and that this occurred more when those losing jobs were white.

<!-- FIGURE X: TIME SERIES OF MANUFACUTRING EMPLOYMENT  -->

@@

```
\begin{figurenotes}
"Gain or loss of national manufacturing jobs, divided by total national workforce."
\end{figurenotes}
\end{figure}

@@

\begin{figure} \label{counterfactual}
\caption{Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center"}

@@

\appendix

\section{Appendix}

<!-- FIGURE X: V SHAPRED RECOVERY  -->

\begin{figure} \label{vShapePlot}
\caption{V-Shaped Recovery in Manufacturing Job Losses}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center"}

ntile_summary <- readRDS("../v_shaped_recovery.rds")

ntile_summary <- ntile_summary %>% mutate(n_tile_04_11 = as.numeric(n_tile_04_11))
# plot 2004-2011 manu layoffs by deciles
ggplot(data = ntile_summary,
       aes(x = n_tile_04_11, y = mean_04_11)) +
  theme_minimal() +
  ylab("Manufacturing Layoffs (04-11)") +
  xlab("Decile")+
  ggtitle("Manufacturing Layoffs from 2004-2011 by Deciles") +
  geom_col(fill = "#4287f5") +
  scale_x_continuous(breaks=seq(1,10,1))

```
\begin{figurenotes}
"Stuff here"
\end{figurenotes}
\end{figure}

\begin{figure} \label{vShapePlot}
\caption{V-Shaped Recovery in Manufacturing Job Losses}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center"}

# plot 2012-2015 manu layoffs by deciles
ggplot(data = ntile_summary,
       aes(x = n_tile_04_11, y = mean_12_15)) +
  theme_minimal() +
  ylab("Manufacturing Layoffs (12-15)") +
  xlab("Decile of Manufacturing Layoffs (04-11)")+
  ggtitle("Manufacturing Layoffs from 2012-2015 by 2004-2011 Deciles") +
  geom_col(fill = "#4287f5") +
  scale_x_continuous(breaks=seq(1,10,1))

\begin{figurenotes}
"Stuff here"
\end{figurenotes}
\end{figure}

### 2021-12-12 Final Draft/Final-Draft.tex
tweaks
pushing changes to final draft

This paper analyzes the relationship between deindustrialization, race,
and the election of Donald Trump in 2016. Previous analysis has
frequently focused on racial animus, trade, or the conjunction of the
two (\cite{Autor20}; \cite{Che16}; \cite{BR21}). We extend the analysis
of \cite{Baccini21}, one of the first papers to examine the political
effects of deindustrialization while considering race and localization.
Deindustrialization - the transition of the U.S. economy away from
manufacturing - has led to widespread job loss and created adverse
knock-on effects for a wide swath of Americans. This incorporates job
losses due to trade with China and others (see \cite{Acemoglu16}); it
also includes job losses due to automation, demand changes, and other
factors. The experience of deindustrialization has not been uniform;
rather, it has been concentrated in certain areas of the country that
were previously economically reliant on manufacturing, as demonstrated
in Figure \ref{ManuMap}. We believe the localized nature of this
analysis highlights the uneven and concentrated harms of
deindustrialization.

Along with \cite{Baccini21}, we find evidence that the experience of
deindustralization was associated with support for Donald Trump,
particularly among whites. However, we find that the related effects in
the 2016 election are not associated with recent manufacturing job loss;
rather, the opposite. Counties with manufacturing job losses from
2012-2015 tended to become more Democratic. We extend the analysis to
the 2004-2015 period, to capture the long-term effects of
deindustrialization on a region. Doing so, we find that regions
experiencing more manufacturing job losses swung towards Donald Trump in
2016, and that this occurred more when those losing jobs were white.

\begin{figure} \label{ManuMap}
\caption{Net Change in Manufacturing Employment from 2004 to 2015}

@@

@@

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-3-1} \end{center}

\begin{figurenotes}
Gain or loss of national manufacturing jobs, divided by total national workforce.
\end{figurenotes}
\end{figure}

\begin{figure} \label{vShapePlot}
\caption{}

\begin{figurenotes}
Figure notes without optional leadin.
\end{figurenotes}
\end{figure}

@@

\begin{figure} \label{natlPlot}
\caption{Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}

@@

>>>>

This paper analyzes the relationship between deindustrialization, race,
and the election of Donald Trump in 2016. Previous analysis has focused
on racial animus, trade, or the conjunction of the two on political
polarization and election outcomes (\cite{Autor20}; \cite{Che16};
\cite{BR21}). We extend the analysis of \cite{Baccini21}, one of the
first papers to examine the political effects of deindustrialization
that while considering race and localization. Deindustrialization - the
transition of the U.S. economy away from manufacturing - has led to
widespread job loss and created adverse knock-on effects for a wide
swath of Americans. Our measure of dindustrialization incorporates job
losses due to trade with China and others (see \cite{Acemoglu16}); it
also includes job losses due to automation, demand changes, and other
factors. The experience of deindustrialization has not been uniform;
rather, it has been concentrated in certain areas of the country that
were previously economically reliant on manufacturing, as demonstrated
in Figure \label{ManuMap}. The localized nature of this analysis
highlights the uneven and concentrated harms of deindustrialization.

\begin{figure} \label{ManuMap}
\caption{Concentration of Deindustrialization (2004-2015) in the eastern Heartland}

@@

Like \cite{Baccini21}, we find evidence that the experience of
deindustralization was associated with support for Donald Trump,
particularly among whites. \cite{Baccini21} find that gross
manufacturing layoffs between 2012-2015 were associated with a loss in
Democratic vote share. However, using a different measure of
deindustrialization - \emph{net} change in manufacturing employment - we
find the oppposite effect: that counties with manufacturing job
\emph{gains} from 2012-2015 tended to swing towards Donald Trump in
2016. We extend the analysis to the 2004-2015 period in order to capture
the long-term effects of deindustrialization on a region. Doing so, we
find that regions experiencing more manufacturing job losses swung
towards Donald Trump in 2016, and that this occurred more when those
losing jobs were white.

@@

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-3-1} \end{center}
\begin{figurenotes}
"Gain or loss of national manufacturing jobs, divided by total national workforce."
\end{figurenotes}
\end{figure}

@@

\begin{figure} \label{counterfactual}
\caption{Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}

@@

\appendix

\section{Appendix}

\begin{figure} \label{vShapePlot}
\caption{V-Shaped Recovery in Manufacturing Job Losses}

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-8-1} \end{center}
\begin{figurenotes}
"Stuff here"
\end{figurenotes}
\end{figure}

\begin{figure} \label{vShapePlot}
\caption{V-Shaped Recovery in Manufacturing Job Losses}

\begin{center}\includegraphics{Final-Draft_files/figure-latex/unnamed-chunk-9-1} \end{center}

\begin{figurenotes}
"Stuff here"
\end{figurenotes}
\end{figure}
